{\Huge Supersymmetry} \\ \vspace{5mm} Lecture notes for FYS5190/FYS9190 Supersymmetry

### 2025-11-10 lagrangian.tex
Some exercise answers

\begin{Exercise}[]
\label{ex:simpleL}
Write down the Lagrangian and find the action of the simplest possible supersymmetric field theory with a single scalar superfield, without gauge transformations, in terms of component fields, and show that it contains no kinetic terms for the $F$-field. Then show how the $F$-field can be eliminated by the equations of motion. {\it Hint:} The kinetic part of the action will turn out to be
\begin{equation}
S_\text{kin}=\int d^4x \left\{-A^*(x)\square A(x)+|F(x)|^2+i\partial_\mu\psi(x)\bar\sigma^\mu\psi(x)\right\}.
\end{equation}
To show this you may have use of the identities in Sec.~\ref{sec:Weylspinor_calc}.
\end{Exercise}

\begin{Answer}
The Lagrangian for a single scalar superfield $\Phi$ is
\begin{equation}
\mathcal{L}=\Phi^\dagger \Phi + \delta^2(\bar{\theta}) (g\Phi+m\Phi\Phi+\lambda\Phi\Phi\Phi) + \delta^2(\theta) (g\Phi^\dagger+m\Phi^\dagger\Phi^\dagger+\lambda\Phi^\dagger\Phi^\dagger\Phi^\dagger). 
\end{equation}

Using the following expressions for the scalar superfield in terms on component fields
\begin{eqnarray}
\Phi(x, \theta, \bar{\theta}) &=& A(x) + i(\theta\sigma^\mu \bar{\theta})\partial_\mu A(x) - \frac{1}{4}\theta\theta\bar{\theta}\bar{\theta}\Box A(x) + \sqrt{2}\theta \psi(x) - \frac{i}{\sqrt{2}}\theta\theta\partial_\mu \psi(x)\sigma^\mu\bar{\theta} \nonumber\\
&&+ \theta\theta F(x).\nonumber\\
\Phi^\dagger (x, \theta, \bar{\theta}) &=& A^*(x) - i(\theta\sigma^\mu \bar{\theta})\partial_\mu A^*(x) - \frac{1}{4}\theta\theta\bar{\theta}\bar{\theta}\Box A^*(x) + \sqrt{2}\bar{\theta}\bar{\psi}(x) + \frac{i}{\sqrt{2}}\bar{\theta}\bar{\theta}\theta \sigma^\mu\partial_\mu \bar{\psi}(x)\nonumber\\
&& +\bar{\theta}\bar{\theta} F^*(x).\nonumber
\end{eqnarray}
we have for the kinetic term,
\begin{eqnarray}
\int d^4\theta \,\Phi^\dagger\Phi&=& - \frac{1}{4}A\Box A^* -\frac{1}{4} A^*\Box A  +  |F|^2 \nonumber \\
&&+\int d^4\theta \left\{ (\theta\sigma^\mu \bar{\theta})\partial_\mu A^*(\theta\sigma^\nu \bar{\theta})\partial_\nu A
+i\bar{\theta}\bar{\theta}\theta \sigma^\mu\partial_\mu \bar{\psi}\theta \psi - i\bar{\theta}r{\psi}\theta\theta\partial_\mu \psi\sigma^\mu\bar{\theta}\right\},\nonumber
\end{eqnarray}
where we have removed the factor $\theta\theta\bar{\theta}\bar{\theta}$ by integration. Using the identities
\begin{eqnarray}
(\theta\sigma^\mu \bar{\theta})(\theta\sigma^\nu \bar{\theta})&=&\frac{1}{2}g^{\mu\nu}\theta\theta\bar{\theta}\bar{\theta},\\
\theta\sigma^\mu\partial_\mu \bar\psi \theta\psi &=& -\frac{1}{2}\psi\sigma^\mu\partial_\mu \bar\psi \theta\theta,\\
\partial_\mu\psi\sigma^\mu \bar\theta \bar\theta\bar\psi &=& -\frac{1}{2}\partial_\mu\psi\sigma^\mu \bar\psi \bar\theta\bar\theta,
\end{eqnarray}
gives
\begin{eqnarray}
\int d^4\theta \,\Phi^\dagger\Phi&=& - \frac{1}{4}A\Box A^* -\frac{1}{4} A^*\Box A +\frac{1}{2}\partial^\mu A\partial_\mu A^* +  |F|^2 \nonumber \\
&&-\frac{i}{2}\psi\sigma^\mu\partial_\mu \bar\psi+ \frac{i}{2}\partial_\mu\psi\sigma^\mu \bar\psi.\nonumber
\end{eqnarray}
Now, since
\begin{equation}
\frac{1}{2}\partial^\mu A\partial_\mu A^* = \frac{1}{2}\partial^\mu (A\partial_\mu A^*)-\frac{1}{2}A\Box A^*,
\end{equation}
and
\begin{eqnarray}
-\frac{1}{4}A\Box A^* &=& -\frac{1}{4}\partial_\mu (A\partial^\mu A^*)+\frac{1}{4}\partial_\mu A\partial^\mu A^*\nonumber\\
&=& -\frac{1}{4}\partial_\mu (A\partial^\mu A^*)+\frac{1}{4}\partial^\mu ((\partial_\mu A) A^*)-\frac{1}{4} A^*\Box A,\nonumber
\end{eqnarray}
we can write
\begin{eqnarray}
- \frac{1}{4}A\Box A^* -\frac{1}{4} A^*\Box A +\frac{1}{2}\partial^\mu A\partial_\mu A^* = -A^*\square A + {\rm total~derivatives.}
\end{eqnarray}
Using (\ref{eq:Weylspinor_etasigmapsi}) we can similarly write
\begin{eqnarray}
-\frac{i}{2}\psi\sigma^\mu\partial_\mu \bar\psi+ \frac{i}{2}\partial_\mu\psi\sigma^\mu \bar\psi
&=&\frac{i}{2}\partial_\mu \bar\psi\bar\sigma^\mu\psi- \frac{i}{2}\bar\psi\bar\sigma^\mu\partial_\mu\psi\nonumber\\
&=&i\partial_\mu \bar\psi\bar\sigma^\mu\psi- \frac{i}{2}\partial_\mu(\bar\psi\bar\sigma^\mu\psi).
\end{eqnarray}
Removing the terms with total derivatives the kinetic term is
\begin{equation}
\int d^4\theta \,\Phi^\dagger\Phi=\int d^4x \left\{-A^*(x)\square A(x)+|F(x)|^2+i\partial_\mu\bar\psi(x)\bar\sigma^\mu\psi(x)\right\}.
\end{equation}

>>>>

\begin{Exercise}[]
\label{ex:simpleL}
Write down the Lagrangian and find the action of the simplest possible (renormalisable) supersymmetric field theory with a single scalar superfield  $\Phi$, without gauge transformations, in terms of component fields, and show that it contains no kinetic terms for the $F$-field. Then show how the $F$-field can be eliminated by the equations of motion. {\it Hint:} The kinetic part of the action will turn out to be
\begin{equation*}
S_\text{kin}=\int d^4x \left\{-A^*(x)\square A(x)+|F(x)|^2+i\partial_\mu\psi(x)\bar\sigma^\mu\psi(x)\right\}.
\end{equation*}
To show this you may have use of the identities in Sec.~\ref{sec:Weylspinor_calc}.
\end{Exercise}

\begin{Answer}
The most general Lagrangian for a single scalar superfield $\Phi$ is
\begin{equation*}
\mathcal{L}=\Phi^\dagger \Phi + \delta^2(\bar{\theta}) (g\Phi+m\Phi\Phi+\lambda\Phi\Phi\Phi) + \delta^2(\theta) (g\Phi^\dagger+m\Phi^\dagger\Phi^\dagger+\lambda\Phi^\dagger\Phi^\dagger\Phi^\dagger). 
\end{equation*}

Using the following expressions for the scalar superfield in terms on component fields
\begin{eqnarray}
\Phi(x, \theta, \bar{\theta}) &=& A(x) + i(\theta\sigma^\mu \bar{\theta})\partial_\mu A(x) - \frac{1}{4}\theta\theta\bar{\theta}\bar{\theta}\Box A(x) + \sqrt{2}\theta \psi(x) - \frac{i}{\sqrt{2}}\theta\theta\partial_\mu \psi(x)\sigma^\mu\bar{\theta} \nonumber\\
&&+ \theta\theta F(x).\nonumber\\
\Phi^\dagger (x, \theta, \bar{\theta}) &=& A^*(x) - i(\theta\sigma^\mu \bar{\theta})\partial_\mu A^*(x) - \frac{1}{4}\theta\theta\bar{\theta}\bar{\theta}\Box A^*(x) + \sqrt{2}\bar{\theta}\bar{\psi}(x) + \frac{i}{\sqrt{2}}\bar{\theta}\bar{\theta}\theta \sigma^\mu\partial_\mu \bar{\psi}(x)\nonumber\\
&& +\bar{\theta}\bar{\theta} F^*(x).\nonumber
\end{eqnarray}
we have for the kinetic term,
\begin{eqnarray}
\int d^4\theta \,\Phi^\dagger\Phi&=& - \frac{1}{4}A\Box A^* -\frac{1}{4} A^*\Box A  +  |F|^2 \nonumber \\
&&+\int d^4\theta \left\{ (\theta\sigma^\mu \bar{\theta})\partial_\mu A^*(\theta\sigma^\nu \bar{\theta})\partial_\nu A
+i\bar{\theta}\bar{\theta}\theta \sigma^\mu\partial_\mu \bar{\psi}\theta \psi - i\bar{\theta}\bar{\psi}\theta\theta\partial_\mu \psi\sigma^\mu\bar{\theta}\right\},\nonumber
\end{eqnarray}
where we have removed the factor $\theta\theta\bar{\theta}\bar{\theta}$ by integration in the first three terms. Using the identities
\begin{eqnarray}
(\theta\sigma^\mu \bar{\theta})(\theta\sigma^\nu \bar{\theta})&=&\frac{1}{2}g^{\mu\nu}\theta\theta\bar{\theta}\bar{\theta},\\
\theta\sigma^\mu\partial_\mu \bar\psi \theta\psi &=& -\frac{1}{2}\psi\sigma^\mu\partial_\mu \bar\psi \theta\theta,\\
\partial_\mu\psi\sigma^\mu \bar\theta \bar\theta\bar\psi &=& -\frac{1}{2}\partial_\mu\psi\sigma^\mu \bar\psi \bar\theta\bar\theta,
\end{eqnarray}
gives
\begin{eqnarray}
\int d^4\theta \,\Phi^\dagger\Phi&=& - \frac{1}{4}A\Box A^* -\frac{1}{4} A^*\Box A +\frac{1}{2}\partial^\mu A\partial_\mu A^* +  |F|^2 -\frac{i}{2}\psi\sigma^\mu\partial_\mu \bar\psi+ \frac{i}{2}\partial_\mu\psi\sigma^\mu \bar\psi.\nonumber
\end{eqnarray}
Now, since
\begin{equation}
\frac{1}{2}\partial^\mu A\partial_\mu A^* = \frac{1}{2}\partial^\mu (A\partial_\mu A^*)-\frac{1}{2}A\Box A^*,
\end{equation}
and
\begin{eqnarray}
-\frac{1}{4}A\Box A^* &=& -\frac{1}{4}\partial_\mu (A\partial^\mu A^*)+\frac{1}{4}\partial_\mu A\partial^\mu A^*\nonumber\\
&=& -\frac{1}{4}\partial_\mu (A\partial^\mu A^*)+\frac{1}{4}\partial^\mu ((\partial_\mu A) A^*)-\frac{1}{4} A^*\Box A,\nonumber
\end{eqnarray}
we can write
\begin{eqnarray}
- \frac{1}{4}A\Box A^* -\frac{1}{4} A^*\Box A +\frac{1}{2}\partial^\mu A\partial_\mu A^* = -A^*\square A + {\rm total~derivatives.}
\end{eqnarray}
Using (\ref{eq:Weylspinor_etasigmapsi}) we can similarly write
\begin{eqnarray}
-\frac{i}{2}\psi\sigma^\mu\partial_\mu \bar\psi+ \frac{i}{2}\partial_\mu\psi\sigma^\mu \bar\psi
&=&\frac{i}{2}\partial_\mu \bar\psi\bar\sigma^\mu\psi- \frac{i}{2}\bar\psi\bar\sigma^\mu\partial_\mu\psi\nonumber\\
&=&i\partial_\mu \bar\psi\bar\sigma^\mu\psi- \frac{i}{2}\partial_\mu(\bar\psi\bar\sigma^\mu\psi).
\end{eqnarray}
Removing the terms with total derivatives the kinetic term is
\begin{equation}
\int d^4\theta \,\Phi^\dagger\Phi=\int d^4x \left\{-A^*(x)\square A(x)+|F(x)|^2+i\partial_\mu\bar\psi(x)\bar\sigma^\mu\psi(x)\right\}.
\end{equation}
